Online {\IterComp}

### 2017-10-21 rebuttal.md

#Review#107A
A.1 The metric is in some ways coarse, but has a strong correlation coefficient to runtime (0.99) and mean absolute error of 3.5% (Fig3). It is enough for iterative compilation (Fig9).

A.2 We used all benchmarks in KDataSets. Some were for the cost model, the others for evaluating our methodology. This standard model
training/evaluation practice. We will make this clear in the paper.

A.3 We will publish the cost model.

#Review#107B

The paper is **not** about *reducing overhead in iterative compilation*. It is about **online** iterative compilation. We will clarify this in the paper.

B.1.Offline iterative compilation is done ahead of time in the development environment. Online iterative compilation is done *live* on users' machines. Offline requires developers to choose representative inputs, online does not. Instead, online optimization targets the actual ways users use the program. Because it is live on users' machines, inputs cannot be run twice. We will clarify this.

B.2.Online iterative compilation has **never** been done before in the general case. Some work has done online optimization where the developer provides a work or throughput metric (e.g. #requests/second assuming each request does the same work). No one has solved the general, automatic case without developer involvement. We are the first in the world to do this. Many papers have suggested it is possible. None have done it and their suggestions would not work. We will clarify this.

Yes, there has been work on input sensitivity. We agree that programs' behavior varies over inputs. That is why online iterative compilation is better, it optimizes according to the inputs the user actual presents. Input sensitive iterative compilation (via multi-versioning etc) is entirely orthogonal to online/offline. Offline requires the developer to choose representative inputs, online does not. Online input sensitive iterative compilation will match the users's inputs better. Until now, online has been impossible. We will clarify this.

The papers, Fursin etal., Ding etal., and Chen etal., are about offline iterative compilation, their approaches do not work in the online case.

That runtimes of programs with different input sets are not directly comparable is not a problem for **offline** iterative compilation, it breaks online iterative compilation. Throughput could be used, but there is no such general metric. Developers would have to provide one (e.g. #requests/second - and ensure each request comprises the same work). Our work is the first general and automatic approach.

IPC has been suggested by other papers as a possible approach to achieve online iterative compilation. None has tried it. We have shown that IPC does not work (Fig2). Our example with NOPs is only the most egregious case, to show logically why IPC is no good. In fact, it is not the cause of IPC's poor correlation in Fig2. Trading high CPI instructions for low CPI instructions (e.g. replace divide by $2^k$ with shift) also shows IPC is no good. Filtering NOPs would not fix the other cases. IPC is only good for architectural optimization. Our observation is back up by paper[17]. We will clarify this and change the example from NOPs.

Our work is not about the *overhead of replay executions*. It is about online iterative compilation where inputs cannot be repeated.

By side effects, we mean changes to external states that may not be repeated (e.g., modifying a database, file system, network activity, interacting with external third-party services). Mitigating these effects would be very cumbersome, if it is possible at all. It is different from the scenario faced by ATLAS and not about caches or noisy environments. We will clarify.

Regarding the distinction between online/offline from Fig4, please see B.1,B.2.

Regarding the shipping of the program to the iterative compilation mechanism: We only mean that the users get the program from somewhere and run it inside the iterative compilation framework. The code needs to be re-compilable, so LLVM bitcode is suitable. Application source code is not required. It is poorly worded, we will fix.

Our work is not meant only for data centers. It is generally applicable. We will clarify.

There do not need to be many alternate binaries stored. It depends on the search strategy. In the smallest case, only one binary exists at any time. Many are produced over time. We will clarify.

    For more on autotuning on the data center, see this paper from CGO16

    Dehao Chen, David Xinliang Li, and Tipp Moseley. 2016. AutoFDO: automatic feedback-directed optimization for warehouse-scale applications.

    The authors claim the proposed strategy is suitable for data center environments. But the evaluation is done on a single Intel system
    with sequential applications.

B.6
We agree that the benchmark suite used does not represent typical data center workloads. Still, we believe that our approach is perfectly suited to data center scenarios and there is no reason it would not work. In a future work, we plan to evaluate our mechanism under such an environment to better support our claim.

    The performance improvement numbers are modest. But this is probably because only 500 sequences are selected out of a possible 200
    factorial sequences in LLVM.
Yes, the performance improvements are modest. We choose a simple, repeatable, random search to allow easy comparison with other metrics. Better searches (e.g.genetic algorithms) often do much better, but the search technique is orthogonal to our work and not our focus.

    What was the basis for choosing the subset of programs from KDataSets? This needs to be clearly justified.

Please see response to A.2 regarding benchmark partitions.

    Minor comments:

    Speedup on X-axes made it difficult for me to interpret the results

Figure 9 reports average speedups over O3 optimization level as a % improvement whereas Figure 1 reports improvement as speedup. % improvement and speedup are not the same. We will clarify.

    Why is the Oracle-RM performing worse than other methods in some instances (e.g., qsort)?

    Section 4 has only one subsection. Consider revising.

    page 10, line 1: correlate -> correlated

We will make all the minor corrections you suggested. Thanks for the feedback.

#Review#107C

    Overall merit
    -------------
    2. Weak reject

    Reviewer expertise
    ------------------
    3. Knowledgeable

    Paper summary ------------- 
    The paper presents a new online iterative optimization strategy. The main new contributions of this work
    are a new work efficiency metrics that makes it possible to compare different program versions (compiled with different optimization
    sessions) across different inputs. The paper also includes an optimization of basic block profiling that is based on the idea of
    removing profiling probes if the amount of work they contribute is low but their execution count might be high.

    Comments for author
    -------------------

    I found the new efficiency metric interesting – making it possible to compare different versions of a program across different inputs
    is a novel contribution. However, some of the discussion around the efficiency metric in the paper was confusing. For example, Section
    2 states that execution time is not correlated to speedup – how is speedup exactly computed here – isn’t it based on execution time?
    Section 3 then introduces the work metrics and shows in Figure 2 that it is correlated to execution time. Earlier we established that
    execution is not correlated with speedup so does that mean the new work metric also is not correlated with speedup? The new work is
    however, positioned to approximate speedup – so these analyses are confusing – it may just be a result of the description in the paper
    but the authors should clarify.

C.1:
Thank you for the feedback, we will clarify this confusion in the paper.

In section 2, we mean that execution time alone does not correlate to the actual speedup, when we compare the program execution time obtained from different inputs, as different inputs would entail different amounts of work. In section 3, we show that work metric obtain from a given input indeed correlates with the unoptimized execution time for *the same input*, regardless of the optimization used to compile the program.

    My main problem with the paper is that it introduces a new online iterative optimization strategy without properly evaluating it. The
    paper offers Figure 9 to show that the final optimization sequence chosen by the new online strategy reaches about 60% of the
    performance that can be achieved with offline iterative optimization.

    This evaluation doesn’t show all that much, it shows me that the proposed new strategy “has the potential to produce good results”.
    This is not enough I also need to know what it takes (i.e., the cost ) of using online optimization versus starting out with a fixed
    version that was produced with less optimally selected input in an offline iterative optimization.

    Specially, to compare online versus offline iterative optimization, I need to know the accumulated performance penalty from running
    suboptimal versions in the online setting, until I find the final optimization strategy. The authors don’t want to get into evaluating
    different search strategies – that is fine – but one could always pick a fixed search strategy that is used in both online and offline
    and then compare. To assess the potential of an online techniques I need to understand how quickly it converges or in other words how
    long the iterative optimization process continues. There is also the questions as to whether there are any advantages over offline
    iterative optimization, once the online iterative optimization sequence concludes – the actual inputs may continue to evolve .

C.2
Regarding the cost of online iterative compilation, we have focused our analysis on the profiling overhead required for guiding the search, but not the time spent on searching for the optimal version. We target scenarios where the program to be optimized will be repeatedly executed, so that the search overhead can be amortised by the improved performance. There are works on using active learning and genetic search to reduce the search overhead, which are orthogonal to our work.

Nonetheless, we will provide results on the search overhead, and refer to those complementary techniques.

    In any realistic setting, one would want to start out with the best statically compiled version  - can this strategy provide any benefits over that scenario?

C.3
Indeed it would interesting to investigate a search scheme that leverages from starting with the best statically compiled version of the program. However, unfortunately we do not have data for this experiment, as we assumed that -O3 would be the initial baseline. We will investigate it in the future.

    We need a comparison to offline iterative optimization to understand whether this approach is viable.

C.4
Prior works on offline iterative compilation are all based on replays or requires the profiling to be performed on the same input. These assumptions are not suitable for the scenarios we target (see also response to B.1). We will, however, provide a comparison against offline iterative compilation.

    The limited evaluation included is further limited by the fact that the benchmarks programs are very small and not representative of realistic datacenter applications.

C.5 Please refer to response B.6 regarding the benchmarks we used.

#Review#107D

    Overall merit
    -------------
    2. Weak reject

    Reviewer expertise
    ------------------
    3. Knowledgeable

    Paper summary
    -------------
    Paper proposes online iterative compilation technique, which is composed of
    multiple components.
    1. Work metric to approximate performance.
    2. Relaxation methods that reduces runtime profiling overhead but penalizes
    estimation accuracy.

    Comments for author

    It is difficult to find what authors are newly contributing on top of
    existing works.
    Authors use work efficiency metric to guide optimization selection. While
    authors claim that they introduce a new "work" metric, this metric does not seem
    to be critical because "work efficiency" metric is what authors actually use to
    guide the optimization selection.
    e.g. if authors defined work efficiency as 1 / t, where t is execution time, it
    would have resulted in same conclusions.
    Although not very clear from the paper, if authors compute "work" metric
    differently for every optimization, then authors are defining the metric wrong,
    as authors highlight that work metric is defined as the amount of time taken by
    the "unoptimized" program.

D.1
Indeed, the work metric does not depend on optimizations but it does depend on the input. More input to process or more complex processing result in more work. Each measurement is done using not only a different program version but also a different input, so the work is different for each measurement. Our work efficiency metric takes this work and the runtime to calculate the amount of work done in a unit of time. Since both of them change, the work efficiency is not similar to the inverse of the runtime.

    The methods described for reducing instrument (or profiling) overhead at runtime
    are also similar to the prior works, and does not seem to contribute any
    novelty.
    In fact, many of the implementation details in the paper could just cite the
    previous work, and authors should have focused on the new parts.

D.2
We will make the contributions more explicit in section 5. As we highlight in the introduction, the main contributions regarding the profiling technique are the two relaxation strategies. Apart from the relaxation strategies, however, the optimal work profiling is not a direct implementation of the classic block-frequency profiling, thus we judged important to also describe it in more detail.

    Another problem is that there are many inconsistencies on authors' claims. For
    example, authors argue that online real inputs are important because they could
    be very different from small set of inputs given by programmer.
    However, some of the methods used by authors' do require those pre-knowledges.
    For example, whole-program relaxation requires block-frequency profiling. Since
    the instrumentation itself is done once at offline, block-frequency profiling
    must be also done before online deployment. This profiling will be again
    dependent on those small set of inputs given by programmer, which contradicts
    what authors were aiming for.

D.3
It is possible to use static heuristics for estimating block frequencies. In fact, these static heuristics are used in the results for the worst-case relaxation, where no prior profiling is required. For the whole-program profiling, however, we want to show the results for the ideal case by using actual block-frequency profiling. These block-frequency profiling can be integrated with the online framework, for example, by interleaving a sequence of executions with one that only performs the block-frequency profiling.

    Minor typos, etc:
    - Page 6, second to last paragraph:
      into the DAG An example of this is
      -> period missing between "DAG" and "An".

    - Page 9, last paragraph:
      user, by it also
      -> "by" should be "but"

### 2017-10-21 rebuttal.md

#Review#107A
A.1 The metric is in some ways coarse, but has a strong correlation coefficient to runtime (0.99) and a mean absolute error of 3.5% (Fig3). It is enough for iterative compilation (Fig9).

A.2 We used all benchmarks in KDataSets. Some were for the cost model, the others for evaluating our methodology. This standard model
training/evaluation practice to evaluate the generality of the model. We will make this clear in the paper.

@@

#Review#107B

The paper is **not** about *reducing overhead in iterative compilation*. It is about **online** iterative compilation. We will clarify this in the paper.

B.1.Offline iterative compilation is done ahead of time in the development environment. Online iterative compilation is done *live* on users' machines. Offline requires developers to choose representative inputs, online does not. Instead, online optimization targets the actual ways users use the program. Because it is live on users' machines, inputs cannot be run twice. We will clarify this.

B.2.Online iterative compilation has **never** been done before in the general case. Some work has done online optimization where the developer provides a work or throughput metric (e.g. #requests/second assuming each request does the same work). No one has solved the general, automatic case without developer involvement. We are the first to do this. We will clarify this.

B.3 Yes, there has been work on input sensitivity. We agree that programs' behavior varies over inputs. That is why online iterative compilation is better, it optimizes according to the inputs the user actual presents. Input sensitive iterative compilation (via multi-versioning etc) is entirely orthogonal to online/offline. Offline requires the developer to choose representative inputs, online does not. Online input sensitive iterative compilation will match the users's inputs better. Until now, online has been impossible. We will clarify this.

@@

The papers, Fursin etal., Ding etal., and Chen etal., are about offline iterative compilation, their approaches do not work in the online case.

That runtimes of programs with different input sets are not directly comparable is not a problem for **offline** iterative compilation, it breaks online iterative compilation. Throughput could be used, but there is no such general metric. Developers would have to provide one (e.g. #requests/second - and ensure each request comprises the same work). Our work is the first general and automatic approach.

B.4 IPC has been suggested by other papers as a possible approach to achieve online iterative compilation. None has tried it. We have shown that IPC does not work (Fig2). Our example with NOPs is only the most egregious case, to show logically why IPC is no good. In fact, it is not the cause of IPC's poor correlation in Fig2. Trading high CPI instructions for low CPI instructions (e.g. replace divide by $2^k$ with shift) also shows IPC is no good. Filtering NOPs would not fix the other cases. IPC is only good for architectural optimization where the program is invariant. Our observation is back up by paper [17]. We will clarify this and change the example from NOPs.

B.5 Our work is not about the *overhead of replay executions*. It is about online iterative compilation where inputs cannot be repeated. See also B.1 and B.2.

B.6 By side effects, we mean changes to external states that may not be repeated (e.g., modifying a database, file system, network activity, interacting with external third-party services). Mitigating these effects would be very cumbersome, if it is possible at all. It is different from the scenario faced by ATLAS and not about caches or noisy environments. We will clarify.

B.7 Regarding the distinction between online/offline from Fig4, please see B.1 and B.2.

B.8 Regarding the shipping of the program to the iterative compilation mechanism: We only mean that the users get the program from somewhere and run it inside the iterative compilation framework. The code needs to be re-compilable, so LLVM bitcode is suitable. Application source code is not required. It is poorly worded, we will fix.

B.9 Our work is not meant only for data centers. It is generally applicable. We will clarify.

B.10 There do not need to be many alternate binaries stored. It depends on the search strategy. In the smallest case, only one binary exists at any time. Many are produced over time. We will clarify.

### 2017-10-21 rebuttal.md

Regarding the shipping of the program to the iterative compilation mechanism: We only mean that the users get the program from somewhere and run it inside the iterative compilation framework. The code needs to be re-compilable, so LLVM bitcode is suitable. Application source code is not required. It is poorly worded, we will fix.

Our work is not meant only for data centers. It is generally applicable. We will clarify.

There do not need to be many alternate binaries stored. It depends on the search strategy. In the smallest case, only one binary exists at any time. Many are produced over time. We will clarify.

The AutoFDO paper is not about iterative compilation. It repeatedly profiles and feeds profiles to an FDO compiler, but this is not iterative compilation search (despite a section with that title) -- theirs has no search for performance efficiency. We will discuss in related work.

Yes, our evaluation was not done in a data center. However, it is entirely general and should work from mobile systems to data centers. We did not mean to suggest that we had evaluated it in all domains. We will clarify.

Yes, the performance improvements are modest because of the small, random search. We choose it to allow easy comparison with other metrics. Better searches (e.g.genetic algorithms) often do much better, but the search technique is orthogonal to our work and not our focus.

@@

Please see response to A.2 regarding benchmark partitions.

Figure 9 reports average speedups over O3 optimization level as a % improvement whereas Figure 1 reports improvement as speedup. % improvement and speedup are not the same. We will clarify.

Regarding Oracle-RM performing worse than other methods, these measurements are within the noise floor.  We will run more times increase the run count to get statistically significant comparisons.

### 2017-10-21 rebuttal.md

A.3 We will publish the cost model.

#Review#107B

The paper is **not** about *reducing overhead in iterative compilation*. It is about **online** iterative compilation. We will clarify this in the paper.

B.1 Offline iterative compilation is done ahead of time in the development environment. Online iterative compilation is done *live* on users' machines. Offline requires developers to choose representative inputs, online does not. Instead, online optimization targets the actual ways users use the program. Because it is live on users' machines, inputs cannot be run twice. We will clarify this.

B.2 Online iterative compilation has **never** been done before in the general case. Some work has done online optimization where the developer provides a work or throughput metric (e.g. #requests/second assuming each request does the same work). No one has solved the general, automatic case without developer involvement. We are the first to do this. We will clarify this.

B.3 Yes, there has been work on input sensitivity. We agree that programs' behavior varies over inputs. That is why online iterative compilation is better, it optimizes according to the inputs the user actual presents. Input sensitive iterative compilation (via multi-versioning etc) is entirely orthogonal to online/offline. Offline requires the developer to choose representative inputs, online does not. Online input sensitive iterative compilation will match the users's inputs better. Until now, online has been impossible. We will clarify this.

@@

The papers, Fursin etal., Ding etal., and Chen etal., are about offline iterative compilation, their approaches do not work in the online case.

That runtimes of programs with different input sets are not directly comparable is not a problem for **offline** iterative compilation, it breaks online iterative compilation. Throughput could be used, but there is no such general metric. Developers would have to provide one (e.g. #requests/second - and ensure each request comprises the same work). Our work is the first general and automatic approach.

B.4 IPC has been suggested by other papers as a possible approach to achieve online iterative compilation. None has tried it. We have shown that IPC does not work (Fig2). Our example with NOPs is only the most egregious case, to show logically why IPC is no good. In fact, it is not the cause of IPC's poor correlation in Fig2. Trading high CPI instructions for low CPI instructions (e.g. replace divide by $2^k$ with shift) also shows IPC is no good. Filtering NOPs would not fix the other cases. IPC is only good for architectural optimization where the program is invariant. Our observation is back up by paper [17]. We will clarify this and change the example from NOPs.

B.5 Our work is not about the *overhead of replay executions*. It is about online iterative compilation where inputs cannot be repeated. See also B.1 and B.2.

B.6 By side effects, we mean changes to external states that may not be repeated (e.g., modifying a database, file system, network activity, interacting with external third-party services). Mitigating these effects would be very cumbersome, if it is possible at all. It is different from the scenario faced by ATLAS and not about caches or noisy environments. We will clarify.

B.7 Regarding the distinction between online/offline from Fig4, please see B.1 and B.2.

B.8 Regarding the shipping of the program to the iterative compilation mechanism: We only mean that the users get the program from somewhere and run it inside the iterative compilation framework. The code needs to be re-compilable, so LLVM bitcode is suitable. Application source code is not required. It is poorly worded, we will fix.

B.9 Our work is not meant only for data centers. It is generally applicable. We will clarify.

### 2017-10-21 rebuttal.md

# Review#107A
A.1 The metric is in some ways coarse, but has a strong correlation coefficient to runtime (0.99) and a mean absolute error of 3.5% (Fig3). It is enough for iterative compilation (Fig9).

@@

Please see response to A.2 regarding benchmark partitions.

Figure 9 reports average speedups over O3 optimization level as a % improvement whereas Figure 1 reports improvement as speedup. % improvement and speedup are not the same. We will clarify.

Regarding Oracle-RM performing worse than other methods, these measurements are within the noise floor.  We will run more times increase the run count to get statistically significant comparisons.

# Review#107C

C.1 Thank you for the feedback, we will clarify this confusion in the paper.

In Section 2 we mean that execution time does not correlate to efficiency unless all inputs have the same work. We will reword.

In Section 3 we show the work metric correlates with unoptimized execution time for *the same input*, regardless of optimization. We will clarify.

We think you are asking for a "total lifetime cost", i.e. including costs of poorly performing program versions in the search. This is a very good idea which we did not think of. However, random search is ill suited for this, a well designed GA would be better. We were not considering the search itself as interesting. We will show total lifetime cost, and in the future apply better search.

Starting with the best statically compiled version is a very good idea. We don't know the quantitive benefit. We will consider this in future work when improving the search technique.

We do compare offline iterative compilation, OracleRM, it runs each input twice per optimization. Our approach is gets close to this.

Regarding benchmarks not being for data centers, please see B.9.

# Review#107D

For novelty, please see B.2.

Yes, work efficiency is the metric. It is work/time. Time is easy, we need a work metric.

If work efficiency were $1/t$: consider comparing efficiency of two sorting algorithms this way. If program A sorts a 2 element list and program B sorts a million, $t_A<t_B$ but we do not know which is more efficient. We will clarify in the paper.

Our work metric does not change because of the optimization. We will clarify.

The only similarity in our work is to classic block-frequency profiling. We extend with relaxation which is entirely novel. We will rebalance the writing.

The profiling that drives the whole program relaxation is done *online* - by profiling the first few inputs (not programmer provided). It could be periodically updated. We regard this as somewhat orthogonal. No pre-knowledge is required. We will clarify.

Thank for catching the typos.

### 2017-10-21 rebuttal.md

The paper is **not** about *reducing overhead in iterative compilation*. It is about **online** iterative compilation. We will clarify this in the paper.

B.1 Offline iterative compilation is done ahead of time in the development environment. Online iterative compilation is done *live* on users' machines. Offline requires developers to choose representative inputs, online does not. Instead, online optimization targets the actual ways users use the program. Because it is live on users' machines, inputs cannot be run twice. We will clarify this.

B.2 Online iterative compilation has **never** been done before in the general case. Some work has done online optimization where the developer provides a work or throughput metric (e.g. #requests/second assuming each request does the same work). No one has solved the general, automatic case without developer involvement. We are the first to do this. We will clarify this.

B.3 Yes, there has been work on input sensitivity. We agree that programs' behavior varies over inputs. That is why online iterative compilation is better, it optimizes according to the inputs the user actual presents. Input sensitive iterative compilation (via multi-versioning etc) is entirely orthogonal to online/offline. Online input sensitive iterative compilation will match the users's inputs better. We will clarify this.

The papers, Fursin etal., Ding etal., and Chen etal., use offline iterative compilation, their approaches do not work in the online case.

That runtimes of programs with different input sets are not directly comparable is not a problem for **offline** iterative compilation, but it breaks online iterative compilation. Throughput could be used, but there is no such general metric. Developers would have to provide one (e.g. #requests/second - and ensure each request comprises the same work). Our work is the first general and automatic approach.

B.4 IPC has been suggested by other papers as a possible approach to achieve online iterative compilation. None has tried it. We have shown that IPC does not work (Fig2). Our example with NOPs is only the most egregious case, to show logically why IPC is no good. In fact, it is not the cause of IPC's poor correlation in Fig2. Trading high CPI instructions for low CPI instructions (e.g. replace divide by $2^k$ with shift) also shows IPC is no good. Filtering NOPs would not fix the other cases. IPC is only good for architectural optimization where the program is invariant. We will clarify this and change the example from NOPs.

@@

Yes, the performance improvements are modest because of the small, random search. We chose it to allow easy comparison with other metrics. Better searches (e.g.genetic algorithms) often do much better, but the search technique is orthogonal to our work and not our focus.

Please see response to A.2 regarding benchmark partitions.

Figure 9 reports average speedups over O3 optimization level as a % improvement whereas Figure 1 reports improvement as speedup. % improvement and speedup are not the same. We will clarify.

Regarding Oracle-RM performing worse than other methods, these measurements are within the noise floor.  We will increase the run count to get statistically significant comparisons.

@@

C.1 Thank you for the feedback, we will clarify this confusion in the paper.

In Section 2 we mean that execution time does not correlate to efficiency unless all inputs have the same work. We will reword.

In Section 3 we show the work metric correlates with unoptimized execution time for *the same input*, regardless of optimization. We will clarify.

We think you are asking for a "total lifetime cost", i.e. including costs of poorly performing program versions in the search. This is a very good idea which we did not think of. However, random search is ill-suited for this, a well designed GA would be better. We were not considering the search itself as interesting. We will show total lifetime cost, and in the future apply better search.

@@

# Review#107D

For novelty, please see B.2.

Yes, work efficiency is the metric. It is work/time. Time is easy, we need a work metric.

**HJL rewrite - lists of different sizes**
If work efficiency were $1/t$: consider comparing efficiency of two sorting algorithms this way. If program A sorts a 2 element list and program B sorts a million, $t_A<t_B$ but we do not know which is more efficient. We will clarify in the paper.

### 2017-11-02 motivation.tex
done with related work
Alternative motivation section. Old one saved in another file

\section{Motivation}

    In this section we first motivate online iterative compilation by showing the deleterious effects of choosing the wrong inputs for the
    offline case. Then we go on to show the unsuitability of previously proposed program version comparison metrics.

    \begin{figure}[t!]
        \centering
        \includegraphics[width=0.5\textwidth]{figs/motivation-online.pdf}
        \caption{
            Histograms of speedup over \texttt{-O3} for each of the 1,000 inputs of \texttt{susan\_c}. The green histogram shows the
            speedup when using the optimization sequence selected by offline \itercomp, the blue one is the speedup using the optimization
            sequence selected by an online-like approach. \FIXME{Beware Daltonism!}
        }
        \label{fig:motivation-online}
    \end{figure}

    We performed offline iterative compilation on \texttt{susan\_c} to find the optimization settings that maximize the average performance
    over five randomly chosen inputs from a set of 1,000 provided by the \textit{KDataSets} benchmark suite~\cite{chen10,chen12a}.
    The green histogram in Figure~\ref{fig:motivation-online} shows for each of the 1,0000 inputs the speedup versus \texttt{-O3} of that
    chosen optimization setting. The dark green line shows the average speedup for all those inputs. The five training inputs are
    all among the top performing inputs. On the other hand, the chosen optimization does poorly on other inputs, often causing
    a slowdown and yielding only a slight speedup of 4\% on average.

    The blue bars in Figure~\ref{fig:motivation-online} show a different approach. Here the iterative compilation proceeds by evaluating
    each optimization on five inputs. Each input is evaluated only once during the search, but the search is provided with oracle
    knowledge about the unoptimized runtime for the inputs, giving a perfect speedup metric. The environment of the search constantly
    changes in the same way as for online iterative compilation. The best optimization for this case never degrades performance compared to
    \texttt{-O3} and is on average 13\% better than the offline approach. The online case, considering truly representative inputs, leads
    to superior performance, if the speedup caused by optimizing can be estimated.

    \begin{figure}[t]
        \centering
        %\begin{tabular}{cc}
        %\includegraphics[width=0.25\textwidth]{figs/motivation-metric-ipc.pdf} &
        %\includegraphics[width=0.25\textwidth]{figs/motivation-metric-runtime.pdf}
        %\end{tabular}
        \includegraphics[width=0.5\textwidth]{figs/motivation-metric.pdf}
        \caption{
            Relationship between IPC and execution time \textit{vs} speedup over the unoptimized program
            for 500 different binaries of \texttt{susan\_c}.
            Each point represents the average over a subset of all inputs.
            The IPC and execution time show correlation coefficients of $-0.14$ and $-0.16$, respectively.
        }
        \label{fig:motivation-metric}
    \end{figure}

    Prior to this paper there were two proposed metrics to correlate with program efficiency. The first is simply to use runtime, ignoring
    the work aspect. This functions well when the inputs all require similar effort. When that is not the case it bears little or no
    correlation to efficiency. The second approach is \textit{instructions per cycle}~(IPC), suggested by~\citep{fursin07}. While IPC seems
    promising in the sense that faster instruction execution is a good thing, this is belied by the simple observation that adding
    redundant NOPs to a program will increase IPC while also slowing the program down. Figure~\ref{fig:motivation-metric} shows the
    relationship between execution time, IPC and speedup for different optimizations of \texttt{susan\_c}, averaged across a subset of the
    inputs. Neither IPC nor execution time correlate well, and so neither can be used for iterative compilation. We need a work efficiency
    metric which can approximate speedup instead, using only a single execution.

    In the remainder of this paper we present our work efficiency metric and demonstrate its utility for iterative compilation.

>>>>

% vim: set tabstop=4 shiftwidth=4 expandtab : 
\section{Motivation \label{sec:motivation}}

    % Representative inputs are important. Example.
    % Artificial representative input sets are hard to create
    % Representative input sets created out of real data are also hard to create
    % Just using real data online is not enough. Efficiency metrics are needed
    % We do not have a proper efficiency metric. Example.
    % The paper shows how to construct one.

    In this section we first motivate online iterative compilation by showing the deleterious effects of using non-representative input
    sets and that it is hard to create representative ones. Then we go on to examine what stops us from just using real inputs online and
    what we need to overcome the problem.

    \paragraph{Iterative compilation needs representative inputs.} \FIXME{Fig.XXX} shows what happens when we apply iterative compilation
    on \FIXME{susan\_c} using the wrong inputs. We selected \FIXME{50} of the 1,000 inputs of \textit{KDataSets}~\cite{chen10,chen12a} for
    \FIXME{susan\_c} and we used them to evaluate different optimizations. The red bars show the average speedup over \texttt{-O3} achieved
    by the optimization sequence selected by iterative compilation. On the left is the average over the \FIXME{50} inputs, on the right the
    average over all 1,000 inputs. The selected sequence works great for the inputs we used, but it fails miserably for the rest, slowing
    down the benchmark by \FIXME{5\%} on average.

    The blue bars, on the other hand, show the ideal case where good optimizations are selected based on their average effect on all
    inputs, not just a subset. Not only is the average speedup over all 1,000 inputs considerable, \FIXME{13\%} over \texttt{-O3}, but it
    also works well for the \FIXME{50} input subset. The speedup for the subset is only \FIXME{1\%} less of what we got when we optimized
    specifically for that subset.

    Making iterative compilation work in this example is not a matter of handling input sensitivity, we were able to find an optimization
    sequence that works well both on average and under most inputs. It is a matter of finding a set of inputs that represents the full
    range of inputs the application is likely to process. If we do, we can improve average performance by \FIXME{13\%}. If not, we might
    hurt average performance instead.

    \paragraph{Representative input sets are hard to create.} For developers to create artificial input sets, they first need to know how
    their application is actually being used. This is not always easy or even possible. Then, they need to create data that will trigger a
    similar behavior and this needs a deep knowledge of the application and considerable effort. This will have to be a continuous work in
    progress, to handle changes in the way the application is used.

    Collecting real input data online and processing them again offline to evaluate optimizations is not a solution either. It requires
    manual modifications to the application to save the data, more modifications to avoid side effects while reusing data, and even more
    modifications to ensure that the application does the same work every time the same data are reused, despite changes in the rest of
    the system state. This is significant engineering work. Even if this was not the case, many applications are just not able to save
    data for privacy or storage overhead reasons.

    \paragraph{Using real inputs online is not enough.} We could overcome these problems by just evaluating optimization sequences online,
    while the application is being used. This gives us real data but creates a new problem. Iterative compilation finds good optimizations
    by measuring the efficiency of the resulting binary in terms of speedup over another baseline binary. This requires processing the same
    input in the exact same way twice, one with the optimized and one with the baseline binary, measuring the runtimes, and calculating the
    speedup. When done online, we cannot process the same input twice, each time we process whatever the input happens to be. We need
    another efficiency metric that only requires a single use of an optimized binary.

    \paragraph{Existing efficiency metrics are not up to the task.} A application-specific ones~\cite{alameldeen06,coppa14} may work well,
    but the developer has to choose the appropriate metric and then manually modify the application to calculate it. Existing
    application-agnostic ones do not work. We run each of the \FIXME{32} \textit{KDataSets} benchmarks 1,000 times, each time with a
    different input \textit{and} a different optimized binary. We selected three performance metrics that can be easily measured in a
    single run and can potentially be used as efficiency metrics: runtime, inverse runtime, and
    \textit{Instructions Per Cycle}~(IPC)~\citep{fursin07}.

    \FIXME{Fig.XXX} shows how well these three candidates for an efficiency metric correlate with the metric we would actually like to have,
    speedup over \texttt{-O0}. For all benchmarks, the correlation coefficient~\cite{Bishop:2006:PRM:1162264} is below \FIXME{0.20} for
    runtime and below \FIXME{0.35} for inverse runtime. Both have obvious reasons for being inappropriate efficiency metrics. Doubling the
    amount of data to be processed will cause the runtime to rise and the inverse runtime to fall, even if we use the same binary.
    Efficiency has nothing to do with it. IPC also correlates poorly with performance speedup, the correlation coefficient being just
    \FIXME{0.15}. While it is a common efficiency metric in computer architecture, it works there because it is used with invariant binaries.
    In contrast, any compiler optimization that trades high latency instructions for multiple low latency instructions can reduce IPC while
    improving efficiency.

    No existing metric can give us an application-agnostic measure of efficiency with a single execution. The rest of the paper will show
    how to create one and how to use it for iterative compilation.
